\textbf{Sharp Bias-variance Tradeoffs of Hard Parameter Sharing in High-dimensional Linear Regression}

### 2020-10-22 arxiv_full_version/setup.tex
arxiv full version
polish typo

\medskip
\noindent\textbf{Main results.}
Our first result (Theorem \ref{thm_many_tasks}) applies to multi-label prediction settings where all datasets have the same features (and sample size), and we want to make several predictions for every input (cf. examples in \citet{hsu2009multi}).
We analyze the global minimizer of $f(A, B)$, and provide a sharp bias-variance decomposition of its (out-of-sample) prediction loss for any task.
This setting is tractable even though in general, $f(A, B)$ is non-convex in $A$ and $B$ (e.g. matrix completion is a special case for suitably designed $X^{(i)}, Y^{(i)}$).
%We show that the prediction loss of HPS admits a clean bias-variance decomposition.
Our result implies that when all tasks have the same features but different labels, for any task, HPS helps reduce the task's variance compared to single-task learning but increases bias.

@@

Finally, we discuss the practical implications of our work.
Our sample size ratio study implies a concrete progressive training procedure that gradually adds more data until performance drops.
For example, in the setting of Figure \ref{fig_intro_sample_size_b}, this procedure will stop right at the minimum of the local basin.
We conduct further studies of this procedure on six text classification datasets and observe that it reduces the computational cost by $65\%$ compared to a standard round-robin training procedure while keeping the average accuracy of all tasks simultaneously.

\medskip
\noindent\textbf{Proof techniques.}
There are two main ideas in our analysis. The proof of our first result uses a geometric intuition that hard parameter sharing finds a ``rank-$r$'' approximation of the datasets.
We carefully keep track of the concentration error between the global minimizer of $f(A, B)$ and its population version (cf. equation \eqref{eq_A_star}).
The proof of our second result is significantly more involved because of different sample sizes and covariate shifts. We show that the inverse of the sum of two sample covariance matrices with arbitrary covariate shifts converges to a deterministic diagonal matrix asymptotically (cf. Theorem \ref{thm_main_RMT}).
We use recently developed techniques from random matrix theory to show a sharp convergence rate.
% to obtain a sharp estimate on the..., which is commonly referred to as the \emph{local law}.
%\HZ{add several sentences on the technical insight}
One limitation of our analysis is that in Example \ref{ex_sample_ratio}, there is an error term that can result in vacuous bounds for very small $n_1$ (cf. equation \eqref{cor_MTL_error}).
We believe our result has provided significant initial insights, and it is an interesting question to tighten our result.

@@

\smallskip
\noindent\textbf{Organizations.}
The rest of this paper is organized as follows.
In Section \ref{sec_same}, we present the bias-variance decomposition for hard parameter sharing.
In Section \ref{sec_diff}, we describe how varying sample sizes and covariate shifts impact hard parameter sharing using random matrix theory.
In Sections \ref{sec_simulation}, we validate our results in simulations.
In Section \ref{sec_conclude}, we summarize our work and discuss future work.
Section \ref{sec_text} describes our study on text classification tasks.
Section \ref{app_proof_error_same_cov}, \ref{appendix RMT}, and \ref{app_iso_cov} present proofs of our results.

\medskip
\noindent\textbf{Notations.}
%Let $\cE \define [\varepsilon_1, \varepsilon_2, \dots, \varepsilon_t] \in \real^{n \times t}$ denote the random noise.
%We can also write $Y = XB^{\star} + \cE$.
%Let $A = [A_1, A_2, \dots, A_t] \in \real^{r\times t}$ be a matrix notation that contains all the output layer parameters.
For an $n\times p$ matrix $X$, let $\lambda_{\min}(X)$ denote its smallest singular value and $\norm{X}$ denote its largest singular value.
Let $\lambda_1(X), \lambda_2(X), \cdots, \lambda_{p\wedge n}(X)$ denote the singular values of $X$ in decreasing order.
Let $X^+$ denote the Moore-Penrose psuedoinverse of $X$.
We refer to random matrices of the form $\frac {X^\top X} n$ as sample covariance matrices.
We say that an event $\Xi$ holds with high probability if the probability that $\Xi$ happens goes to $1$ as $p$ goes to infinity.
%We shall use $\oo(1)$ to mean a small positive quantity that converges to 0 as $p$ goes to infinity.

>>>>

\medskip
\noindent\textbf{Main results.}
Our first result (Theorem \ref{thm_many_tasks}) applies to multi-label prediction settings where all datasets have the same features (and sample size), and we want to make several predictions for every input (e.g. \citet{hsu2009multi}).
We analyze the global minimizer of $f(A, B)$, and provide a sharp bias-variance decomposition of its (out-of-sample) prediction loss for any task.
This setting is tractable even though in general, $f(A, B)$ is non-convex in $A$ and $B$ (e.g. matrix completion is a special case for suitably designed $X^{(i)}, Y^{(i)}$).
%We show that the prediction loss of HPS admits a clean bias-variance decomposition.
Our result implies that when all tasks have the same features but different labels, for any task, HPS helps reduce the task's variance compared to single-task learning but increases bias.

@@

Finally, we discuss the practical implications of our work.
Our sample size ratio study implies a concrete progressive training procedure that gradually adds more data until performance drops.
For example, in the setting of Figure \ref{fig_intro_sample_size_b}, this procedure will stop right at the minimum of the local basin.
We conduct further studies of this procedure on six text classification datasets and observe that it reduces the computational cost by more than half compared to a standard round-robin training procedure while keeping the average accuracy of all tasks simultaneously.

\medskip
\noindent\textbf{Proof techniques.}
There are two main ideas in our analysis. The proof of our first result uses a geometric intuition that hard parameter sharing finds a ``rank-$r$'' approximation of the datasets.
We carefully keep track of the concentration error between the global minimizer of $f(A, B)$ and its population version (cf. equation \eqref{eq_A_star}).
The proof of our second result is significantly more involved because of different sample sizes and covariate shifts. We show that the inverse of the sum of two sample covariance matrices with arbitrary covariate shifts converges to a deterministic diagonal matrix asymptotically (cf. Theorem \ref{thm_main_RMT}).
We use recently developed techniques from random matrix theory to show a sharp convergence rate.
% to obtain a sharp estimate on the..., which is commonly referred to as the \emph{local law}.
%\HZ{add several sentences on the technical insight}
One limitation of our analysis is that in Example \ref{ex_sample_ratio}, there is an error term that can result in vacuous bounds for very small $n_1$ (cf. equation \eqref{cor_MTL_error}).
We believe our result has provided significant initial insights, and it is an interesting question to tighten our result.
See Section \ref{sec_conclude} for more discussions of the technical challenge.

@@

\smallskip
\noindent\textbf{Organizations.}
The rest of this paper is organized as follows.
In Section \ref{sec_same}, we present the bias-variance decomposition for hard parameter sharing.
In Section \ref{sec_diff}, we describe how varying sample sizes and covariate shifts impact hard parameter sharing using random matrix theory.
In Sections \ref{sec_exp}, we validate our results in simulations and conduct further studies on text classification tasks.
In Section \ref{sec_conclude}, we summarize our work and discuss future work.
Section \ref{app_proof_error_same_cov} presents the missing proofs of Section \ref{sec_same}.
Section \ref{appendix RMT} and \ref{app_iso_cov} show the missing proofs of Section \ref{sec_diff}.

\medskip
\noindent\textbf{Notations.}
%Let $\cE \define [\varepsilon_1, \varepsilon_2, \dots, \varepsilon_t] \in \real^{n \times t}$ denote the random noise.
%We can also write $Y = XB^{\star} + \cE$.
%Let $A = [A_1, A_2, \dots, A_t] \in \real^{r\times t}$ be a matrix notation that contains all the output layer parameters.
For an $n\times p$ matrix $X$, let $\lambda_{\min}(X)$ denote its smallest singular value and $\norm{X}$ denote its largest singular value.
Let $\lambda_1(X), \lambda_2(X), \cdots, \lambda_{p\wedge n}(X)$ denote the singular values of $X$ in decreasing order.
Let $X^+$ denote the Moore-Penrose psuedoinverse of $X$.
We refer to random matrices of the form $\frac {X^\top X} n$ as sample covariance matrices.
We say that an event $\Xi$ holds with high probability if the probability that $\Xi$ happens goes to $1$ as $p$ goes to infinity.
We use the big-O notation $g(n) = \OO(f(n))$ if there exists a constant $C$ such that $g(n) \le C  f(n)$ for large enough $n$.
%We shall use $\oo(1)$ to mean a small positive quantity that converges to 0 as $p$ goes to infinity.
